### 2017-02-02 chapters/ch2-checkpoint.tex
Checkpoint consensus, some definitions

\section{Preliminaries}

\subsection{Requirements}
\begin{itemize}
\item Permissionless
\item Byzantine fault tolerant
\item No PoW
\item Works under churn
\item Underlying data structure is TrustChain
\item Detects forks or double-spends
\item No step in the protocol blocks transactions
\item Application independent
\end{itemize}

\subsection{Assumptions}
\begin{itemize}
\item We elect $N$ consensus promoters in every round, we assume the number of
  faulty promoters is $f$ and $N = 3f + 1$.
\item Promoters have the complete history of the previously agreed set of
  transactions.
% \item Single connected component for TrustChain.
% \item No sybil attack % 
\end{itemize}

\subsection{Notations}
\begin{itemize}

\item $y = H(x)$ is a cryptographically secure hash function (random oracle),
  the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.

\item Every node in the system has an identifier $i$ and a blockchain
  $$B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$$
  where $h$ is the height of the chain. Note that $h = |B_i|$

\item Each block $b_{i,j}$ has a type $t \in \{ \tau, \gamma \}$, denoted by
  $b^t_{i,j}$. Blocks without the superscript can be of any type.

\item $T(b_{i, j}) = \{ \tau, \gamma \}$ is the type function, where its domain
  is a block and outputs the corresponding type of the block.

\item A block of type $\tau$ is a \emph{transaction block}. It is a six-tuple,
  i.e. $$b^\tau_{i,j} = (H(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$
  denote the height (the sequence number for when the transaction is made) of
  the sender and receiver respectively. $s_s$ and $s_r$ denote the signature of
  the sender and the receiver respectively.

\item A block of type $\gamma$ is a \emph{checkpoint block}. It is a
  three-tuple, i.e.
  $$b^\gamma_{i,j} = (H(b_{i,j}), H(\mathcal{C}_r), p)$$
  where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
  which indicates whether $i$ wish to become a promoter in the following
  consensus round.

\item Given the input $b^\tau_{i,j}$, we define the
  \emph{get-neighbouring-checkpoints function}
  $$C(b^\tau_{i,j}) = (b^\gamma_{i,a}, b^\gamma_{i,b})$$
  where $a = \argmin_{k, k < j, T(b_{i,k}) = \gamma}(j - k)$ and $b =
  \argmin_{k, k > j, T(b_{i,k}) = \gamma}(k - j)$.

\item Given two $\gamma$ transactions, we define \emph{get-piece function}
  $$P(b^\gamma_{i,a}, b^\gamma_{i,b}) = \{b_{i,j} : b_{i,j} \in B_i, a \le j \le
  b\}.$$

\item Given two blockchains and a message, we define the \emph{do-transaction
    function}
  $$X_\tau(B_s, B_r, m) = (B'_s, B'_r)$$
  where $B'_s = \{(H(b_{s,h_s}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_s$,
  $B'_r = \{(H(b_{r, h_r}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_r$, $h_s =
  |B_s|$ and $h_r = |B_r|$.

\item Given a blockchain, we define the \emph{do-checkpoint function}
  $$X_\gamma(B_i, r, p) = \{ (H(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$$ where
  $h = |B_i|$, and $\mathcal{C}_r$ is the latest consensus result.

\item Note that $X_\tau$ and $X_\gamma$ perform a state transition.

\item The result of a consensus in round $r$ is a set of two-tuple of
  checkpoints. Namely, $\mathcal{C}_r = \{ (b^\gamma_{i,a}, b^\gamma_{i,b}) : a
  < b, \text{agreed by the promoters}\}$.
\end{itemize}

\section{Checkpoint consensus}

\subsection{Promoter registration}
Node $i$ can register as a promoter when the latest consensus result is
announced (suppose after the completion of round $r-1$), then it generates a new
block using $b = T_\gamma(B_i, r-1, 1)$. The current promoters (in round $r$)
may decide to include $b$ in the new consensus result. If $b$ is in it, then $i$
becomes one of the promoter of round $r+1$.

We can fix the number of promotors to $N$ by sorting the promotors by their
``luck value''' and taking the first $N$.

\subsection{Setup phase}
We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
been disseminated yet.
\begin{lemma}
  Nodes in the network can always verify the validity of the consensus result.
\end{lemma}
\begin{proof}
\end{proof}

\begin{lemma}
  The new set of promoter for the next consensus round is consistent with
  respect to all the nodes in the network.
\end{lemma}

\begin{lemma}
  Promoters waiting for a some time $\Delta$ to collect transactions does not
  violate the asynchronous assumption.
\end{lemma}

\begin{corollary}
  The setup phase satisfies the validity, correctness and termination
  properties.
\end{corollary}

\subsection{Consensus phase}

\section{Fraud detection}
Here we provide two techniques for fraud detection. The first guarantees fraud
detection but is not practical. The second is a randomised solution that detects
fraud with a high probability.

\subsection{Breadth first search}

\subsection{Random sampling}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: