### 2017-02-10 chapters/ch2-checkpoint.tex
Improve definition and write consensus protocol

\section{Preliminaries}

\subsection{Requirements}
\begin{itemize}
\item Permissionless
\item Byzantine fault tolerant
\item No PoW
\item Works under churn
\item Underlying data structure is TrustChain
\item Detects forks or double-spends
\item No step in the protocol blocks transactions
\item Application independent
\end{itemize}

\subsection{Assumptions}
\begin{itemize}
\item Asynchronous network
\item Private and authenticated channel
\item We elect $N$ consensus promoters in every round, we assume the number of
  faulty promoters is $f$ and $N = 3f + 1$.
\item Promoters have the complete history of the previously agreed set of
  transactions (TX).
% \item Single connected component for TrustChain.
% \item No sybil attack % 
\end{itemize}

\subsection{Notation and definition}
\begin{itemize}

\item $y = \texttt{h}(x)$ is a cryptographically secure hash function (random
  oracle), the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.

\item We model our system in the permissionless setting, where each
  participating party has a unique identity $i$, and a chain $B_i$.

\item A chain is a collection of blocks
  $B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$
  the blocks are linked together using hash pointers, similar to bitcoin. All
  blocks contain a reference to the previous block, the very first block with no
  references is the genesis block.

  The sequence number of the block begins at 0 on the genesis block and is
  incremented for every new block. The height of the chain is $h = |B_i|$.

\item There are two sets of blocks $T_i$ and $C_i$, where $T_i \cup
  C_i = B_i$ and $T_i \cap C_i = \emptyset$. This can be seen as the
  block type, where $t_{i,j}$ and $c_{i,j}$ to represent a
  \emph{transaction block (TX block)} and \emph{checkpoint block (CP block)}
  respectively.

\item $\texttt{typeof}: B_i \rightarrow \{ \tau, \gamma \}$ returns the
  corresponding type of the block.

\item A block of type $\tau$ is a six-tuple, i.e. $$t_{i,j} =
  (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$ denote the
  height (the sequence number for when the TX is made) of the sender and
  receiver respectively. $s_s$ and $s_r$ are the signatures of the sender and
  the receiver respectively. $i = \{s, r\}$ and $j = \{h_s, h_r\}$ depending on
  whether $i$ is the sender or the receiver.

\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, but $i \neq i'$. If $h_s =
  h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then we call them
  a \emph{pair}.

\item If there exist two TX blocks $t_{i,j}$ and $t_{i',j'}$, where $s_s$ and
  $s_s'$ is created by the same public key, $h_s = h_s'$, but $i \neq i'$, then
  we call this a \emph{fork}.

\item A block of type $\gamma$ is a three-tuple, i.e.
  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), p)$$
  where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
  which indicates whether $i$ wish to become a promoter in the following
  consensus round.

\item We define
  $$\texttt{newtx}: B_i \times B_j \times M \rightarrow B_i \times B_j$$
  as a function that creates new TX blocks. Its functionality is to
  extend the given chains using the following rule. If the input is $(B_s, B_r,
  m)$ then the output is $(B'_s, B'_r)$ where $B'_s = \{(\texttt{h}(b_{s,h_s}), h_s + 1,
  h_r + 1, s_s, s_r, m)\} \cup B_s$, $B'_r = \{(\texttt{h}(b_{r, h_r}), h_s + 1, h_r + 1,
  s_s, s_r, m)\} \cup B_r$, $h_s = |B_s|$ and $h_r = |B_r|$.

\item We define
  $$\texttt{newcp}: B_i \times \mathbb{R}_{\ge 1} \times \{0, 1\} \rightarrow B_i$$
  as a function that creates new CP blocks. Concretely, given $(B_i, r, p)$, it
  results in $\{ (\texttt{h}(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$ where $h =
  |B_i|$, and $\mathcal{C}_r$ is the latest consensus result at round $r$.

\item Note that in the actual system, \texttt{newtx} and \texttt{newcp} perform
  a state transition.

\item The result of a consensus in round $r$ is a set of two-tuple of CP.
  Namely, $\mathcal{C}_r = \{ (c_{i,j}, c_{i,k}) : j < k, \text{agreed by the
    promoters}\}$.
\end{itemize}

\subsection{Properties}
\begin{itemize}
\item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
  $t_{i',j'}$, then $(b_{i, j-1}, b_{i', j'-1})$ is the input block of $t_{i,j}$.

\item A CP block has one \emph{ancestor} block, that is simply the block with
  the previous sequence number, i.e. the ancestor of $c_{i,j}$ is $b_{i,j-1}$.

\item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
  c_{i,b})$, where $a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j
  - k)$ and $b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j)$.

\item The blocks between two $\gamma$ blocks $(c_{i,a}, c_{i,b})$ is $\{b_{i,j}
  : b_{i,j} \in B_i, a \le j \le b\}$.

\end{itemize}

\section{Checkpoint consensus}

\subsection{Luck value}
First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$ is the
public key of $i$. A lower luck value equates to higher luck.

\subsection{Promoter registration}
Node $i$ can register as a promoter when the latest consensus result is
announced (suppose after the completion of round $r-1$), then it generates a new
block $b_{i,j} = \texttt{newcp}(B_i, r-1, 1)$. The current promoters (in round
$r$) may decide to include $b_{i,j}$ in the new consensus result. If $b$ is in
it, then $i$ becomes one of the promoter of round $r+1$.

We can fix the number of promoters to $N$ by sorting the promoters by their
``luck value''' and taking the first $N$.

\subsection{Setup phase}
The setup phase should satisfy the BFT conditions regarding the promoter
selection, that is:
\begin{enumerate}
\item \emph{Agreement}: If any correct node outputs a promoter $p$, then every
  correct node outputs $p$.
\item \emph{Total Order}: If one correct node outputs the sequence of promoters
  $\{p_1, p_i, \dots, p_n\}$ and another has output $\{p'_0, p'_1, \dots,
  p'_{n'} \}$, then $p_i = p'_i$ for $i \le \min(n, n')$.
\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
\end{enumerate}

We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
been disseminated yet. The exact technique to disseminate $\mathcal{C}_{r-1}$ is
irrelevant, broadcasting or gossiping are both sufficient. In fact,
dissemination is not necessary, nodes interested in the result can simply query
the promoters that created $\mathcal{C}_{r-1}$.

\begin{lemma}
  If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
  $\mathcal{C}'_{r-1}$, then $\mathcal{C}_{r-1} = \mathcal{C}'_{r-1}$.
\end{lemma}
\begin{proof}
  (sketch) $\mathcal{C}_{r-1}$ is signed by at least $N-f$ promoters from round
  $r-1$.
\end{proof}

The potential promoters now need to first discover whether
they are the first $N$ lucky promoters.

\begin{lemma}
  The new set of promoter for the next consensus round is consistent with
  respect to all the nodes in the network.
\end{lemma}
\begin{proof}
  (sketch) All nodes use the same deterministic function to compute the luck
  value.
\end{proof}

Nodes should wait for some time to collect the CP blocks, so they wait for some
time $\Delta$ before moving on to the next phase.

\begin{lemma}
  Promoters waiting for a some time $\Delta$ to collect transactions does not
  violate the asynchronous assumption.
\end{lemma}
\begin{proof}
  (sketch) This can be seen as a long delay in the asynchronous system.
\end{proof}

\begin{corollary}
  The setup phase satisfies the validity, agreement and liveness properties.
\end{corollary}
\begin{proof}
  (sketch) Validity and agreement is satisfied because the promoters are
  computed using a deterministic algorithm using the latest consensus result.
  Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
  all node by gossiping.
\end{proof}

\subsection{Consensus phase}
The consensus phase should satisfy the BFT conditions regarding the CP blocks,
that is:
\begin{enumerate}
\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
  correct node outputs $c$.
\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
\end{enumerate}

We need an atomic broadcast algorithm for the consensus phase. We use a similar
but simplified construction as~\cite{miller2016honey}, where atomic broadcast is
constructed from the reliable broadcast\footnote{Reliable broadcast solves the
  Byzantine generals problem.}~\cite{bracha1984asynchronous} and asynchronous
common subset (ACS). The ACS protocol requires a binary Byzantine agreement
protocol, and for that it needs a trusted dealer to distributed the secret
shares. Promoters can check whether the secret shares are valid, but they cannot
prevent the dealer from disclosing the secrets.

There techniques that uses no dealers. First is to use
PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
into the weak synchrony assumption. Most likely this is not possible because of
the ``When to start?'' problem. It's difficult to give a bounded delay for
propagating the consensus result.

Second is to use an inefficient binary Byzantine agreement protocol where its
message complexity is $O(N^3)$ rather than $O(N^2)$ and becomes a bottleneck. Or
a suboptimal one, e.g. $n/5$ instead of $n/3$.

Suppose we use a dealer, what is the effect to the algorithm if the dealer is
malicious?

\section{Fraud detection}

\subsection{Threat model}
\begin{itemize}
\item What is the influence of a fork for future transactions?
\item What if both parties (sender and receiver) are malicious?
\end{itemize}

\subsection{Random sampling}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End: