### 2017-03-15 presentation.tex
Presentation of current work

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{graphicx}

\title{Presentation Title}

% A subtitle is optional and this may be deleted
\subtitle{Optional Subtitle}

\author{K.~Cong}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Delft University of Technology] % (optional, but mostly needed)
{
  Faculty of Electrical Engineering, Mathematics and Computer Science\\
  Delft University of Technology}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

% \date{Conference Name, 2013}
\date{2017}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

% \subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Background}

\subsection{TrustChain}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-good}
  \centering
  \caption{\emph{TX block} is a six-tuple: $t_{i,j} = (\texttt{h}(b_{i,j-1}),
    h_s, h_r, s_s, s_r, m)$, one transaction results in two TX blocks---\emph{a
      pair}.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad}
  \centering
  \caption{Fork is two correctly signed TX blocks that has the same $h_s$ but
    involve different receivers. Only one TX block may be in consensus.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}
  \begin{itemize}
    \item Everyone has their own chain
    \item Transactions are on arbitrary data $m$
    \item Transactions make the chains intertwined
    \item Transactions are irrefutable due to hash pointers
    \item No consensus (my thesis)
  \end{itemize}
\end{frame}

\section{My Thesis}
\subsection{TrustChain with Checkpoints}
\begin{frame}{TrustChain with Checkpoints}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad-cp}
  \centering
  \caption{CP block is a five-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
    \texttt{h}(\mathcal{C}_r), r, p, s)$, $\mathcal{C}_r$ is the consensus
    result at round $r$, $p =$ promoter indicator, $s =$ signature.}
  \end{figure}

\end{frame}

\subsection{Protocol Overview}
\begin{frame}{Protocol Overview}
  \begin{enumerate}
    \item $n$ lucky nodes are selected at random to act as promoters.
    \item Promoters run a BFT (Byzantine Fault Tolerant) consensus algorithm to
      agree on a set of CP blocks.
    \item Disseminate the consensus result (the CP blocks).
    \item Repeat for next round.
    \item Any interested node can validate that their transaction.
  \end{enumerate}
\end{frame}

\subsection{Promoter Registration}
\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-1}
  \centering
  \caption{We start in the state where $\mathcal{C}_{r - 1}$ has just been agreed but
    not yet propagated.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-2}
  \centering
  \caption{Nodes receive consensus result and set promoters indicator $p$, then
    send the new CP blocks to promoters of round $r$.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-3}
  \centering
  \caption{Transactions carry on as usual in round $r$. Note that our CP blocks
    (round $r - 1$) has not reached consensus yet.}
  \end{figure}

\end{frame}

\begin{frame}{Promoter Registration}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-4}
  \centering
  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$. If we're
    lucky: $\texttt{h}(k_i || c_{i,j}) < T$, then we're responsible for
    consensus of round $r + 1$.}
  \end{figure}

\end{frame}

\subsection{Consensus}
\begin{frame}{Consensus}
  \begin{enumerate}
    \item Nodes send CP blocks to the promoters.
    \item The promoters' identities are encoded in the consensus result.
    \item Promoters run the some asynchronous BFT consensus algorithm to agree on
      a set of CP blocks---$\mathcal{C}_r$.
    \item $n = 3t + 1$ is the optimal for BFT consensus.
    \item $\mathcal{C}_r$ and the signatures are disseminated.
    \item Nodes create new CP blocks when they receive $t + 1$ good signatures and
      $\mathcal{C}_r$.
  \end{enumerate}

  \begin{theorem}
    Assuming the promoters satisfy $n = 3t + 1$. The promoter registration and
    the consensus protocol satisfied agreement, total order and liveness.
  \end{theorem}
\end{frame}

\subsection{Validation}

\begin{frame}{Validation}
Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
Suppose $u$ wish to validate $t_{i,j}$. It performs the following.

\begin{enumerate}
\item Determine the pair $t_{i', j'}$.
\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
  $\mathcal{C}_r$, otherwise return ``unknown''.
%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
  correct. Otherwise return ``unknown''.
\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
  Otherwise return ``invalid''.
\item Return ``valid''.
\end{enumerate}
\end{frame}

\begin{frame}{Validation}
  In essence, given a TX, ask the receiver to proof that it has a set of
  transactions that produces some CP block, the CP block should be in the
  consensus result and the pair of the TX should be in that set.

  \begin{theorem}
    If at least one party of every transaction is honest, then forking and other
    forms of tampering is guaranteed to be detected if the malicious party is
    alive.
  \end{theorem}
\end{frame}

\subsection{Implementation}
\begin{frame}{Implementation}
  \begin{itemize}
    \item Currently on going---using Python and Twisted
    \item Completed BFT consensus algorithm
    \item Completed local TrustChain
    \item Next step is networked TrustChain and validation protocol
  \end{itemize}
\end{frame}
\end{document}