### 2017-05-29 thesis/appendices/consensus-example.tex
First draft of protocol overview

\chapter{Consensus Example}
\label{app:consensus-example}

\begin{figure}[htb]
    \frame{\includegraphics[trim={0 0 0 11cm}, clip, width=1.0\textwidth]{trustchain-1}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-1}
\end{figure}

\begin{figure}[htb]
    \frame{\includegraphics[trim={0 0 0 8cm}, clip, width=1.0\textwidth]{trustchain-2}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-2}
\end{figure}

\begin{figure}
    \frame{\includegraphics[trim={0 0 0 2cm}, clip, width=1.0\textwidth]{trustchain-3}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-3}
\end{figure}

\begin{figure}
    \frame{\includegraphics[width=1.0\textwidth]{trustchain-4}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-4}
\end{figure}