### 2017-06-12 thesis/chapters/ch3-analysis.tex
More proofs

\chapter{Analysis}
\label{ch:analysis}

Up to this point we described our system specification in detail.
Of course, specification along does not establish any truths.
In this chapter, we prove two aspects of our system.
First is correctness,
where we show that the consensus protocol and the validation protocol satisfies their desired properties
(\Cref{def:consensus} and \Cref{def:validation} respectively).
The second is performance,
where we prove the lower bound of our throughput and show that it out performs classical blockchain systems.

\section{Correctness}
Our first objective in this section is to establish truths regarding the correctness of our protocol.
We do this in two parts.
First we use mathematical induction to show that properties in \Cref{def:consensus} holds for all round.
Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.

\subsection{Correctness of Consensus}

We begin our analysis by establishing the fact that the $\textsf{get\_facilitator}(\cdot)$ is fair.
\begin{lemma}
\label{lemma:fairness}
\textbf{\emph{Fairness of facilitator election}}

Every node with a CP block in $\C_r$, should have an equal probability to be elected as a facilitator.
\end{lemma}
\begin{proof}
This directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_r, || pk_i)$.
Since $pk_i$ is unique for every node that has a CP block in $\C_r$, the output of $\textsf{H}(\cdot)$ is uniformly random.
This implies that the ordered sequence by luck value is uniformly random.
\end{proof}

Using \Cref{lemma:fairness}, we show that our consensus protocol satisfied \Cref{def:consensus}.
\begin{lemma}
\label{lemma:consensus}
\textbf{\emph{Correctness of consensus}}

$\forall r \in \mathbb{N}$, agreement, validity, fairness and termination holds.
\end{lemma}

\begin{proof}
We proof by mathematical induction.

In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol, due to the bootstrap oracle.
Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1 and are responsible for driving the ACS protocol in round 2.

For the inductive step, we assume that the properties hold for round $r$.
Then, to start round $r + 1$, the honest nodes begin sending CP blocks to $\F_r$. % TODO but they don't send at the same time!
Since the honest region in $\F_r$ waits for $D$ and $D \gg \Delta$,
the CP blocks of the honest nodes are guaranteed to be received by at least $n - t$ facilitators.
The agreement property of ACS (from \Cref{def:acs}) ensures that the $\C_{r + 1}$ is in consensus. 
Observe that $\F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
Thus agreement of $\F_{r + 1}$ follows directly from the agreement of ACS.
The validity property of ACS ensures that $\C_{r + 1}$ contains the input of at least $n - 2t$ parties, but this is a quorum containing at least one honest facilitator,
thus $\C_{r + 1}$ contains the CP blocks of all nodes.
Due to the assumption that the adversary cannot corrupt more than $t$ nodes,
validity of $\F_{r + 1}$ also follows from validity of ACS.
Since all honest nodes are in $\C_{r + 1}$, the fairness property follows directly from \Cref{lemma:fairness}.
Finally, the termination property holds because $D$ eventually elapses and then ACS eventually terminates
(the totality property of ACS).
This completes the proof.
\end{proof}

\subsection{Correctness of Validation}
The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
In this section we build on top of \Cref{lemma:consensus} to show that most properties in \Cref{def:validation} can be satisfied.

\begin{lemma}
\textbf{\emph{Correctness of the validation protocol}}

The validation protocol outputs the correct result
according to the aforementioned validity definition.
\end{lemma}
\begin{proof}
Follows directly from the definition.
\end{proof}

\begin{lemma}
\label{lemma:validation-agreement}
\textbf{\emph{Agreement of the validation protocol}}

If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
\end{lemma}
\begin{proof}
We proof by contradiction.
Without loss of generality, for some transaction $t$ with an agreed fragment $F$,
node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
Then there exist a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
In both cases, the $\textsf{get\_validity}(\cdot)$ must have reached \Cref{line:valid-fragment},
meaning that $c' = c''$, otherwise the result would be \emph{unknown}.
Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
But we assumed that the adversary can only perform polynomial-time algorithm, but in order to find $t''$ it needs to query the random oracle exponentially many times.
Thus we have a contradiction and this completes our proof.
\end{proof}

A consequence \Cref{lemma:validation-agreement} is unforgeability.
That is, no polynomial time adversary is able to create two chains $F = \{ \dots, t, c\}$ $F' = \{ \dots, t', c\}$ with correct hash pointers and the same end of chain $c$.

Now we show a negative result, where the liveness property cannot be attained.
\begin{lemma}
\textbf{\emph{Impossibility of liveness of the validation protocol}}

There exist a valid transactions that cannot be validated eventually.
\end{lemma}
\begin{proof}
We proof by providing a counterexample.
Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction $t$.
Then when $u$ wants to validate $t$, it does so by sending \texttt{vd\_req} message to $v$.
$v$ can act maliciously and ignore all \texttt{vd\_req} message, thus $t$ can never be validated.
\end{proof}
Although this is a negative result, it does not put the adversary in an advantageous position.
If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
Thus, to stay relevant in the system, the adversary need to comply to the protocol.

\subsection{Summary}
We have shown that

TODO: chain structure cannot be satisfied, but can be probabilistically checked

\section{Performance}