### 2017-06-17 thesis/chapters/ch5-related-work.tex
Write some related work

\chapter{Related Work}
\label{ch:related}

Having analysed our system both theoretically and experimentally,
we dedicate this chapter on comparing our results with related work.
Blockchain technology has seen a surge in recent years from both the industry and academia.
There are many approaches and innovation.
We classify by their consensus approach and divide them into the following categories: 
\begin{enumerate}
    \item classical blockchain systems,
    \item Byzantine fault tolerance consensus based consensus,
    \item hybrid systems and
    \item blockchain without global consensus.
\end{enumerate}

\section{Classical Blockchain Systems}
This category represent systems with a probabilistic probabilistic consensus algorithm.
That is, transactions never reach consensus with a probability of 1.
The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
In Bitcoin, the level of consensus of a block\footnote{Note that a block in Bitcoin contain many transactions whereas our TX block only contain a single transaction.}
is determined how deep it is in the Bitcoin blockchain, also called the number of confirmation.
The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
Nevertheless, the probability of the highest block being orphaned is non-negligible.
The advantage of this type of consensus is that it can be used in a large network and is reasonably secure.
Attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
The disadvantage however is that transactions are never in consensus with a probability of 1.
Also, the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.

Our system significantly improves upon Bitcoin and other classical blockchain systems in the performance, scalability and consistency department.
The results described in \Cref{ch:analysis} and \Cref{ch:implementation}, show that we have horizontal scalability,
where more nodes result in more global throughput.
Further, we do not have the aforementioned probabilistic behaviour,
once some consensus result is decided, it cannot be orphaned.
The leadership election is also not ideal in classical blockchain systems.
Mining can be seen as a technique to elect a single leader.
The leader has full control of what goes into the block thus it may selectively censor transactions.
We use ACS, so as long as the CP block is in $n - 2t$ nodes, it is guaranteed to be in consensus.

However, the security aspect falls short of the ``honest majority'' security model that classical blockchains claim to 
have\footnote{Recently it was shown that doing selfish-ming would give the adversary an unfair advantage when she only controls 25\% of the mining power~\cite{eyal2014majority}}.
Our system risks going into erroneous state if the inequality $n \ge 3t + 1$ is not satisfied.
Classical blockchain systems also have an incentive mechanism, thus they do not depend on altruistic nodes and encourages participation.
Our system on the other hand does not have an incentive mechanism because we make no assumptions on the application.

\section{Byzantine Fault Tolerance Consensus Based Systems}

The Byzantine fault tolerance problem is a classical problem is distributed systems literature.
It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.