### 2017-06-21 thesis/chapters/ch3-analysis.tex
Rewrite proof

\chapter{Analysis of Correctness and Performance}
\label{ch:analysis}

Up to this point we described our system specification in detail.
Of course, specification along does not establish any truths.
In this chapter, we analyse two aspects of our system.
First we show it has the desired properties.
That is, the properties in \Cref{def:consensus} and \Cref{def:validation} should hold.
Then we analyse the performance, especially the throughput,
and show that it out performs classical blockchain systems.

\section{Correctness in the Presense of Faults}
Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
Then, building on top of it, we show \Cref{def:validation} holds for the validation protocol.
The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks,
in other words, implicit consensus.

% Our first objective in this section is to establish truths regarding the correctness of our protocol.
% We do this in two parts.
% First we use mathematical induction to show that properties in \Cref{def:consensus} holds for all round.
% Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.

\subsection{Analysis of the Consensus Protocol}

We begin our analysis by establishing truths on the four properties in \Cref{def:consensus},
namely agreement, validity, fairness and termination.
Using these results, we use mathematical induction to show that they hold for all rounds.

\begin{lemma}
\label{lemma:agreement}
For an arbitrary round $r$
if $\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
then all correct nodes output $\F_{r+1}$.
\end{lemma}
\begin{proof}
The argument follows from the protocol description.
Given that $\F_r$ is known,
correct nodes will send CP blocks to all members in $\F_r$.
The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
It cannot make progress until $n-t$ honest facilitators start algorithm,
but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
At the end of ACS, some $\C_{r+1}$ is created, and is broadcasted along with the signature of the facilitators.
Due to the agreement property of ACS (\Cref{def:acs}),
every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
\end{proof}

\begin{lemma}
\label{lemma:validity}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$,
(b) $\F_{r+1}$ must contain at least $n - t$ honest nodes and
(c) $|\F_{r+1}| = n$.
\end{lemma}
\begin{proof}
The validity follows from the validity property of ACS and the definition of our model,
namely $N \ge n + t$ and $n \ge 3t + 1$.
Given $\F_r$, since $N \ge n + t$ , there is at least $n$ nodes that would send their CP block to $\F_r$.
From the validity property of ACS, we know the output must contain the input of at least $n - 2t$ nodes.
But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$, this proves (a).
There are $n-t$ honest nodes in $\F_{r+1}$ follows from the model, this proves (b).
Finally, since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, $|\F_{r+1}| = n$ and this proves (c).
\end{proof}

\begin{lemma}
\label{lemma:fairness}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes then
every node with a CP block in $\C_{r+1}$, should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
\end{lemma}
\begin{proof}
We have already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
Then the proof directly follows from the random oracle model.
Recall that the luck value is computed using $\textsf{H}(\C_{r+1}, || pk_u)$.
Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
the output of $\textsf{H}(\cdot)$ is uniformly random.
This effectively generates a random permutation
so every node has the same probability of being in the top $n$ for the ordered sequence,
namely the output of $\textsf{get\_facilitators}(\cdot)$.
\end{proof}

\begin{lemma}
\label{lemma:termination}
For an arbitrary round $r$,
if $\F_r$ is known by all correct nodes then
every correct node eventually outputs some $\F_{r + 1}$.
\end{lemma}
\begin{proof}
This follows directly from the properties of the channel (eventual delivery)
and the termination property of ACS.
That is, $\F_r$ eventually receives all the CP blocks required to begin ACS.
ACS eventually terminates.
Finally the results are eventually dissemminated to all the nodes.
\end{proof}

From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existance of some $\F_r$.
Thus, to proof the whole of \Cref{def:consensus},
we need to proof these 4 properties under the universal quantifier.
We do this using mathematical induction.

\begin{theorem}
\label{theorem:consensus}
For all rounds,
the consensus protocol satisfies agreement, validity, fairness and termination (\Cref{def:consensus}).
\end{theorem}
\begin{proof}
We proof using mathematical induction.

In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol,
due to the bootstrap oracle.
Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1, who are responsible for driving the ACS protocol in round 2.

For the inductive step,
we assume that the 4 properties hold in round $r$ and prove that they also hold in round $r + 1$.
Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
it directly follows from Modus Poneus that these properties hold for $r + 1$.
Due to the principals of mathematical induction, these properties hold for all $r$.
\end{proof}

\subsection{Correctness of Validation}
The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
In this section we build on top of \Cref{theorem:consensus} to show that most (except liveness) properties in \Cref{def:validation} can be satisfied.

\begin{lemma}
% (Correctness of the validation protocol)
The validation protocol outputs the correct result
according to the validity definition.
\end{lemma}
\begin{proof}
The algorithm (\Cref{alg:get-validity}) is the validity definition.
\end{proof}

\begin{theorem}
\label{theorem:validation-agreement}
% (Agreement of the validation protocol)
If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
\end{theorem}
\begin{proof}
We proof by contradiction.
Without loss of generality, for some transaction $t$ with an agreed fragment $F$,
node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
Then there exist a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
In both cases, the $\textsf{get\_validity}(\cdot)$ function must have reached \Cref{line:valid-fragment}.
Due to \Cref{theorem:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
In other words, whoever sent $F''$ must be able to create some $t''$ that has the same digest as $t'$.
But we assumed that the adversary can only perform polynomial-time algorithm,
so in order to find $t''$ it needs to query the random oracle exponentially many times.
Thus we have a contradiction and this completes the proof.
\end{proof}

\Cref{theorem:validation-agreement} is our first major result.
It shows that consensus on CP blocks would lead to consensus on TX blocks when the nodes are running the validation protocol.
Just like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
One of the main advantages over running a consensus algorithm on all the transactions is that 
the rate of transaction is no longer dependent on the consensus algorithm---ACS.
This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
In addition, a convenient consequence \Cref{theorem:validation-agreement} is unforgeability.
That is, no polynomial time adversary is able to create two chains $F = \{ \dots, t, c\}$ $F' = \{ \dots, t', c\}$ with correct hash pointers and the same end of chain $c$.
% Another way to think about it is if a transaction can be forged, then it cannot be agreed.

\subsection{Impossibility of Liveness}
While \Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
it is not perfect.
Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
Meaning that trasanctions with adversaries cannot always be validated.

\begin{lemma}
There exist a valid transactions that cannot be validated eventually.
\end{lemma}
\begin{proof}
We proof by providing a counterexample.
Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction $t$.
Then when $u$ wants to validate $t$, it does so by sending \texttt{vd\_req} message to $v$.
$v$ can act maliciously and ignore all \texttt{vd\_req} message, thus $t$ can never be validated.
\end{proof}
Although this is a negative result, it does not put the adversary in an advantageous position.
If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
Thus, to stay relevant in the system, the adversary need to comply to the protocol.

% \subsection{Chain structure}
% In this section we discuss how the adversary may tamper with the chain structure and its effects.
% The adversary can create invalid blocks, which are blocks that have an invalid hash pointer.
% The result is some chain with a broken link.

\section{Performance}
TODO intro

\subsection{Message Complexity of ACS}
\label{sec:acs-complexity}
The message complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
where $|v|$ is the size of largest message and $\lambda$ is the security parameter.
Note that the security parameter is the same as the one for our random oracle described in \Cref{sec:model-assumptions}.
In particular, it is from the use of $\textsf{H}(\cdot)$ in the reliable broadcast phase in ACS.
In our system, we wish to understand the scalability properties.
Thus we consider the complexity as a function of $N$ rather than $n$ or $\lambda$.
Since $|v|$ is at most all the CP blocks from every node, we have $|v| = N$.
Therefore the message complexity of ACS in our system is $O(N)$.
Since we use a constant $n$, $O(N)$ message complexity also holds for a single facilitator.

% We argue that with a $O(N)$ message complexity we also have $O(N)$ time complexity.
% It is know that the running time is $O(\log n)$~\cite{miller2016honey}, which translates to $O(1)$ if our $n$ is constant.
% On the other hand, in many distributed algorithms, computational cost is small when compared to the communication cost.
% But in our system, we are more interested in the effect of the message size.

\subsection{Bandwidth Requirement for Transactions}
To make and validate a transaction, the bandwidth required is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment,
the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then $l = (D_{\text{acs}} + D) \cdot r_{\text{tx}}$, where $D_{\text{acs}}$ is the duration an instance of ACS.
But from \Cref{sec:acs-complexity}, we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
% TODO how does message complexity translate to time complexity?
% TODO computational task minimal compared to messages, thus the sending of message is expensive?
This is natural because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer.
This behaviour is also verified experimentally in \Cref{ch:implementation}.

\subsection{Global Throughput}
Suppose every node has some fixed bandwidth capacity $C$ (unit of communication per second),
they make transactions at $r_{\text{tx}}$ per second.
Then we have the inequality $C \ge r_{\text{tx}} l$, where $l$ is the length of the the agree fragment as before.
Rearranging, we get $\frac{C}{l} \ge r_{\text{tx}}$.
With this, we consider two cases, first is  when all the nodes are running at maximum capacity.
Recall that $l$ is of $O(N)$, so as the the population increases, 
the transaction rate must decrease in order to maintain the inverse relationship,
thus $r_\text{tx}$ is of $O(N^{-1})$. % TODO is this formal? use omega?
Therefore, the global throughput would be $O(N^{-1})N = O(1)$.
In the second case, nodes are not running at maximum capacity and $r_{\text{tx}}$ is maintained.
Therefore, if every node runs at $r_{\text{tx}}$,
the global throughput becomes $O(N)$ until $N$ is too large and we go back to the first case.

The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
Then we maintain a constant global throughput.
This result falls a bit short of what we envisioned in the introduction.
However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
In practice, nodes are able to cache agreed fragments.
For instance, if $u$ and $v$ make $x$ transactions in a single round,
then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.

\section{Effect of A Highly Adverserial Environment}