### 2017-07-03 thesis/appendices/additional-graphs.tex
New graphs

\chapter{Additional Graphs}
\label{app:additional-graphs}

\begin{figure}[h]
  \centering
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-duration-vs-population}
    \caption{Transactions are with fixed neighbours}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/consensus-duration-vs-population}
    \caption{Transactions are with random neighbours}
  \end{subfigure}
  \caption{Communication cost for consensus}
  \label{fig:consensus-duration}
\end{figure}