### 2017-07-18 thesis/chapters/ch7-conclusion.tex
Write synchroniser

\chapter{Conclusion}
\label{ch:conclusion}

\section{Future work}