### 2017-07-26 presentation/presentation.tex
Update presentation

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{amsfonts}
\usepackage{graphicx}
\graphicspath{{../figures/}}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\E}{\mathcal{E}}

\algnewcommand\algorithmicupon{\textbf{Upon}}
\algnewcommand\Upon{\item[\algorithmicupon]}

\algnewcommand\algorithmicsend{\textbf{Send}}
\algnewcommand\Send{\item[\algorithmicsend]}

\title{Blockchain Consensus Protocol with Horizontal Scalability}

% A subtitle is optional and this may be deleted
% \subtitle{Optional Subtitle}

\author{K.~Cong}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Delft University of Technology] % (optional, but mostly needed)
{
  Faculty of Electrical Engineering, Mathematics and Computer Science\\
  Delft University of Technology}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

% \date{Conference Name, 2013}
\date{\today}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

% \subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[] % Do nothing for \subsection*
{
  % do nothing
}

\AtBeginSection[] % Do nothing for \section*
{
  \begin{frame}<beamer>
  \frametitle{Outline}
  \tableofcontents[currentsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}
\subsection{Motivation}
\begin{frame}{\subsecname}{}
  \begin{itemize}
    \item Blockchain systems offer an alternative to central authorities for the first time
    \item Market cap and trade volume figures indicate they are here to stay
    \item Early blockchain systems are not scalable (7 TX/s for Bitcoin)
    \item Parameter tuning leads to centralisation
  \end{itemize}
\end{frame}

\subsection{Research question}
\begin{frame}{\subsecname}
    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
    \emph{scalable} and can reach \emph{global consensus?}
\end{frame}

\subsection{Intuition}
\begin{frame}{\subsecname}{}
  \begin{itemize}
    \item A restaurant owner does not report all of its transactions with an central authority
    \item Ocassionally a customer may leave without paying and this event is reported to a central authority
    \item Our blockchain system achieves scalability using the same idea
  \end{itemize}
\end{frame}

\section{System architecture}
\subsection{System model}
\begin{frame}{\subsecname}
  \begin{itemize}
    \item Population size is $N$
    \item $n$ nodes are facilitators, $t$ nodes are malicious (Byzantine)
    \item $n \ge 3t + 1$
    \item $N \ge n + t$
    \item Purely asynchronous channels with eventual delivery
    \item Public key infrastructure
    \item Random oracle model
    \item Application neutural
  \end{itemize}
\end{frame}

\subsection{Architecture overview}
\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{architecture}
  \centering
  \end{figure}
\end{frame}

\subsection{Extended TrustChain}
\begin{frame}{\subsecname}
  \begin{itemize}
    \item Everyone has their own chain
    \item Two types of blocks, transaction (TX) blocks and checkpoint (CP) blocks
    \item A transaction involves two parties and results in two TX blocks
    \item A CP block captures the chain state
    \item TX and CP blocks are chained together using hash pointers
  \end{itemize}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
  \centering
  \caption{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$.
    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$.}
  \end{figure}
\end{frame}

\subsection{Consensus protocol}
\begin{frame}{\subsecname overview}
  \begin{enumerate}
    \item In round $r$, $n$ out of $N$ lucky nodes are selected at random to act as facilitators
    \item Facilitators run a BFT (Byzantine Fault Tolerant) consensus algorithm to
      agree on a set of CP blocks
    \item Disseminate the consensus result $\C_r$, i.e. the CP blocks
    \item Repeat
  \end{enumerate}
\end{frame}

\begin{frame}{\subsecname properties}
\label{def:consensus}
$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{itemize}
    \item \emph{Agreement}:
        If one correct node outputs a list of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{itemize}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-1}
  \centering
  \caption{Suppose we are in a state where $\mathcal{C}_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-2}
  \centering
  \caption{Nodes receive consensus result $\mathcal{C}_{r - 1}$,
    compute the new facilitators $\mathcal{F}_{r-1}$,
    and send out the new CP blocks.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-3}
  \centering
  \caption{Transactions carry on as usual in round $r$,
  while facilitators are trying to reach consensus on the new CP blocks concurrently.}
  \end{figure}
\end{frame}

\begin{frame}{\subsecname}
  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{trustchain-4}
  \centering
  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$.
  The first $n$ nodes ordered by $\textsf{H}(\mathcal{C}_r || pk)$ become $\mathcal{F}_r$.}
  \end{figure}
\end{frame}

\subsection{Transaction protocol}
\begin{frame}{\subsecname}
\begin{itemize}
\item Request (\texttt{tx\_req}) and response (\texttt{tx\_resp}) protocol
\item Two TX blocks containing the same $txid$ are generated
\item Non-blocking
\end{itemize}
\end{frame}

\begin{frame}{\subsecname}
Create $t_{u, h}$ and send $\langle \texttt{tx\_req}, t_{u, h} \rangle$ to start a transaction.
\vspace{5 mm}
\begin{algorithmic}
    \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
    \State $\textsf{new\_tx}(pk_u, m, txid)$ \Comment create and store $t_{u, i}$
    \State store $t_{v, j}$ as the pair of $t_{u, h}$
    \State send $\langle \texttt{tx\_resp}, t_{u, h} \rangle$ to $v$

    \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$  
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
    \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
\end{algorithmic}
\end{frame}

\subsection{Validation protocol}
\begin{frame}{\subsecname overview}
\begin{itemize}
\item Request (\texttt{vd\_req}) and response (\texttt{vd\_resp}) protocol
\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}
\item The goal is to identify which state a given transaction is in
\end{itemize}
\end{frame}

\begin{frame}{\subsecname properties}
\begin{itemize}
    \item \emph{Correctness}:
        The validation protocol outputs the correct result
        according to the aforementioned validity definition.
    \item \emph{Agreement}:
        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
    \item \emph{Liveness}:
        Any valid transactions can be validated eventually.
\end{itemize}
\end{frame}

\begin{frame}{Validity definition}
\textbf{Function} $\textsf{get\_validity}()$ validates the transaction $t_{u, i}$
\vspace{5 mm}
\begin{algorithmic}
    \State Check that $v$ sent the correct agreed fragment,
    otherwise return \emph{unknown}.

    \State
    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
    \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
        \State \Return \emph{invalid}
    \EndIf
    \State \Comment TX exists

    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
    \If{$m \ne m' \vee pk_u \ne pk'_u$}
        \State \Return \emph{invalid}
    \EndIf
    \State \Comment no tampering
    \State \Return \emph{valid}
\end{algorithmic}
\end{frame}

\begin{frame}{\subsecname}
Send $\langle \texttt{vd\_req}, txid \rangle$ to $v$ to begin.
\vspace{5 mm}
\begin{algorithmic}
    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$

    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
        \State $t_{u, i} \gets \text{the transaction identified by } txid$
        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
\end{algorithmic}
\end{frame}

\section{Analysis of correctness and performance}

\subsection{Correctness of the consensus protocol}
\begin{frame}{\subsecname}
\begin{theorem}
For all rounds,
the consensus protocol satisfies agreement, validity, fairness and termination.
\end{theorem}
\begin{proof}(sketch)
Because consensus result are eventually delivered and the properties of ACS,
we get agreement, validity and termination.
Fairness is from the fact that we model $\textsf{H}(\cdot)$ as a random oracle (RO) and the input to the RO is different for every node,
thus the list of nodes ordered by $\textsf{H}(\C_r || pk )$ is a random permutation of those nodes.
\end{proof}
\end{frame}

\subsection{Correctness of the validation protocol}
\begin{frame}{\subsecname}
\begin{theorem}
If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
\end{theorem}
\begin{proof}(sketch)
Proof by contradiction.
For this attack to work, the adversary must be able to create two different fragments but with the same checkpoint enclosure.
We model $\textsf{H}(\cdot)$ as a RO, so the adversary need to query the RO a exponential\footnote{In terms of the security parameter.} number of times.
But the adversary can only query the RO a polynomial number of times.
\end{proof}
\end{frame}

\subsection{Linear global throughput argument}
\begin{frame}{\subsecname}
\begin{itemize}
  \item Throughput has a notion of time but our model is asynchronous
  \item Additional assumptions needed to make the argument---every unit of communication takes a non-negligible amount of time to process
  \item Bandwidth relation---$NC \ge r_{\text{tx}} l$, where $l$ is $O(N)$
  \item If $r_\text{tx}$ satisfies the inequality, then LHS and RHS grows at the same rate,
  thus we have linear global throughput
\end{itemize}
\end{frame}

\section{Experimental results}
\begin{frame}{Implementation and experiment setup}
  \begin{itemize}
    \item Prototype implementation on Github\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}
    \item SHA256 for hash functions and Ed25519 for digital signature
    \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
  \end{itemize}
\end{frame}

\begin{frame}{Throughput vs population size (random neighbour)}
  \begin{figure}[h]
  \includegraphics[width=1.0\textwidth]{neighbour-random/throughput-vs-population}
  \centering
  \end{figure}
\end{frame}

\begin{frame}{Throughput vs population size (fixed neighbour)}
  \begin{figure}[h]
  \includegraphics[width=1.0\textwidth]{neighbour-fixed/throughput-vs-population}
  \centering
  \end{figure}
\end{frame}

\section{Conclusion}
\begin{frame}{\secname}
    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
    \emph{scalable} and can reach \emph{global consensus?}
\begin{itemize}
  \item[\checkmark] Fault tolerant up to $t$ nodes
  \item[\checkmark] Horizontal scalability
  \item[\checkmark] Global consensus
\end{itemize}
\end{frame}

\begin{frame}{Future work}
\begin{itemize}
\item Improve fault tolerance
\item Improve fork detection
\item Analyse the system in the permissionless environment
\item Concrete application
\end{itemize}
\end{frame}

\end{document}