### 2017-08-09 paper/s1-intro.tex
Write/fix paper

\section{Introduction}

% problem description
% research question
% contribution
% organisation