### 2017-08-20 thesis/formalities/preface.tex

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
Since early 2013, the time when I mined my first Bitcoin, 
I became cautious of the security and privacy issues surrounding centralised systems,
and became interested in building a decentralised society.
Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.
Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
This limitation, as well as my personal ambitions, led me to this thesis,
with the goal of creating a truly horizontally scalable solution.

\vspace{1\baselineskip}

\noindent
There are a number of people who have supported me in this endeavour who I would like to thank.
My supervisor Dr. Johan Pouwelse, your unrivalled enthusiasm and guidance have kept me motivated and productive.
Dr. Zhijie Ren, this work would not have been possible without your invaluable input and the numerous discussions which we have had.
Prof. Dick Epema, you taught me a lot on report writing and the use (and misuse) of mathematical formalism.
Members of the Tribler team---Martijn de Vos, Quinten Stokkink and Chengxin Ma,
you have my gratitude for helping me with the implementation and proofreading my work,
and of course the free coffee.
Members of the Cyber Security group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,
I thank both of you for your feedback on my presentations.

\vspace{1\baselineskip}

\noindent
Kelong Cong

\vspace{1\baselineskip}

\noindent
Delft, The Netherlands

\noindent
\today