### 2017-10-31 paper/s6-conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this work, we described \textsc{Checo}, an application neutral blockchain system with horizontal scalability.
Our novel data structure allows nodes to efficiently store transactions and record state using CP blocks.
The round based consensus protocol uses ACS as a building block to reach consensus on CP blocks.
The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
To make transactions, nodes use the simple and non-blocking transaction protocol.
Finally, we introduce a validation protocol which ensures that if an agreed fragment for some transaction exists,
then nodes reach agreement on the validity of that transaction.
The novelty of \textsc{Checo} is that it decouples consensus and transaction validation,
which enables the desirable horizontal scalability property, without employing sharding.

We achieve the properties described in~\Cref{sec:description}.
Namely, our protocol achieves agreement on transactions as we argued in~\Cref{sec:correctness-of-validity}.
Validity is achieved because honest nodes run the $\textsf{get\_validity}(\cdot)$ function, which, in fact, is the validity definition.
Further, the horizontal scalability is demonstrated in~\Cref{sec:implementation},
in the ideal case as well as the worst case.

In the future, we hope to apply our system to a concrete application and evaluate its performance.
Furthermore, we plan to explore a few useful design alternatives such as open membership, total fork detection and fair facilitator election.